Tracy Widom on CENTC distance matrix
========================================================

load("~/PCoA_x.RData")



x$values[,1]
eigenvalues <- x$values[,1]
plot(x$values)

summary(x$vectors)
plot(x$vectors[,1],x$vectors[,2],ylim=c(-.08,.1),xlim=c(-.1,.1))
plot(x$vectors[,1],x$vectors[,2],pch=16,col=rgb(0,0,0,.3),cex=0.4,ylim=c(-.1,.1),xlim=c(-.1,.1),xlab="PCoA 1",ylab="PCoA 2")

length(x$vectors[,1])
##Vince says use library instead, read up on the differences

## Require stuff
```{r}
require("mclust")
require("ape")
require("RMTstat")
```

## New Tracy Widom Plan - PCoA and then TW
## Add into the PCoA,

```{r}
data <- read.csv("Original_centc_DNADIST.csv",row.names=1)
x <- pcoa(data)
x$values[,1]
eigenvalues <- x$values[,1]
biplot(eigenvalues,plot.axes = c(1,2), dir.axis1=1, dir.axis2=1, rn=NULL,)
plot(eigenvalues)
par(mfrow=c(2,2))
plot(ptw(eigenvalues), main="distribution function")
plot(dtw(eigenvalues), main="Density")
plot(qtw(eigenvalues), main="Quantile")
plot(rtw(eigenvalues), main="Randomly Devious")


made_data <- Mmake(data)
TW<-TWcalc(made_data,150) 
km<-which(TW[[2]]>0.05)[1]

```

Cluster individuals based on first km significant principle coordinates
```{r}
pcx<-Mclust(pc$x[,1:km],G=(km+1)) # originally used VVI model
mainC<-pcx$classification
```

Plot stuff (if we can)
```{r}
col.regions=rainbow(km+1)
require(rgl)
plot3d(pc$x[,1],pc$x[,2],pc$x[,3],xlab="PC1",ylab="PC2",zlab="PC3",col=col.regions[mainC],type="s",radius=0.1) 
plot(pc$x[,1]~pc$x[,3])
```


## Below Here is Jeff work from before

## Functions needed for Tracy Widom

### Mmake function (only needed for non-distance data)
```{r}
Mmake<-function(data){
  pcdat<-data
  mv<-apply(pcdat,2,mean,na.rm=TRUE)
	#make vector of sqrt(p*(1-p))
	vv<-mv/2
	vv<-sqrt(vv*(1-vv))  # need to use sd of data, not p(1-p)
	#redo data and vectors to take out monomophic sites
	#dat<-pcdat[,vv>0&is.na(vv)==FALSE] #don't need either but need something named dat below
	#mv<-mv[vv>0&is.na(vv)==FALSE] #we don't need?
	#vv<-vv[vv>0&is.na(vv)==FALSE] #we don't need?
	M<-t((t(pcdat)-mv)/vv)
	M[is.na(M)]<-0
	return(M)	
}
```

### TWcalc function (SM says ignore, as this is specific for SNP data)
```{r}
TWcalc<-function(dat,k){
  x<-dat
	y<-x%*%t(x)
	eig<-eigen(y)
	eig<-eig$values
	#eig<-eig[1:(length(eig))]  #Why do this?
  #n<-ncol(x)  # UNUSED?
	m0=nrow(x)
  
  #I think this is TW residuals
	TWres_<-c()
	for (i in 1:k){
		m=(m0)-i+1
		#m=m
		eiv<-eig[i:(m-1)]
		#n_=n
		n_=((m+1)*sum(eiv)^2)/((m-1)*sum(eiv^2)-sum(eiv)^2)
		S=((sqrt(n_-1)+sqrt(m)))/n_*(1/sqrt(n_-1)+1/sqrt(m))^(1/3)
		u_=(sqrt(n_-1)+sqrt(m))^2/n_
		L1<-(m-1)*eiv[1]/sum(eiv)
		TW_<-(L1-u_)/S
		TWres_<-c(TWres_,TW_) # why cat these? 
	}
	
	#TWres_
	pres<-c()
	for (i in 1:length(TWres_)){
		TW<-TWres_[i]
		dif<-abs(test[,1]-TW)
		p<-test[dif==min(dif),2]
		pres<-c(pres,p)  # why cat these? 
	}
	#pres
	res<-list(TWres_,pres)
	return(res)
}
```

### Do some test data

Read in some test data: JC distance matrix original 218 CENTC from genbank and Tracy-Widom table
(need to fix this so don't have to specify path)
```{r}
#read in table
dist_mat<-read.csv("~/src/tw_centc/test_matrix.csv",row.names=1,header=T)
test<-read.table("twtable.txt", header=FALSE)
```

Standardize matrix and do PC <- don't do this for distance matrix, only for individual observations (i.e. SNP data, phenotypes, environments, etc.)
```{r}
made_x<-Mmake(dist_mat)
```

PCOA from ape package. Check to make sure this is giving us eigenvalues 
```{r}
#pc<-prcomp(made_x) # for nondistance data  
pc<-pcoa(dist_mat) # get details on how this works
```

Skree plot (check to make sure in order)
```{r}
plot(pc$values[,1])
```

Do Tracy Widom. k is number of PCs to evaluate. Here, trying 150.
This TW is specific for SNP data from Patterson et al. 2006 (according to SM).  Need to redo using TracyWidom distribution from RMTstat package. Check details. 
```{r}
TW<-TWcalc(made_x,150) 
km<-which(TW[[2]]>0.05)[1]
```

Cluster individuals based on first km significant principle coordinates
```{r}
pcx<-Mclust(pc$x[,1:km],G=(km+1)) # originally used VVI model
mainC<-pcx$classification
```

Plot stuff (if we can)
```{r}
col.regions=rainbow(km+1)
require(rgl)
plot3d(pc$x[,1],pc$x[,2],pc$x[,3],xlab="PC1",ylab="PC2",zlab="PC3",col=col.regions[mainC],type="s",radius=0.1) 
plot(pc$x[,1]~pc$x[,3])

#the below fails w/o column names correctly
#dim<-1:3
#text3d(pc$x[,dim],text=rownames(M),col=cols,cex=0.6)
```

